---
title: "Group 2: Initial Analysis"
output: html_notebook
Authors: "Blessed Akoma (0773239), Anmol Singh (769066), 
Vikas Sharma (763571), Jagjit Singh (763569),Aman preet Kaur (773401)"
---

###### The dataset used are the results from a marketing campaign run by a bank in Portugal. The purpose of the campaign was to get bank customers to sign up for a term deposit, a banking product they offer.

```{r Loading Libraries}
library(tidyverse)
library(ggplot2)
library(dplyr)
install.packages("gmodels")
install.packages("party")
library(gmodels)
library(party)


bank <- read.csv("C:/Users/ACER/OneDrive/Desktop/hello/bank-additional-full.csv", header = TRUE, sep = ";")
```

```{r Looking at unique values}

#Unique Jobs
unique(bank$job)

#Unique Contact type
unique(bank$contact)

#Unique Education
unique(bank$education)

#Unique Marital Status
unique(bank$marital)

#Unique Housing
unique(bank$housing)

#Unique Default
unique(bank$default)

```


```{r View Attribute Names}
names(bank)

#Rename Column 'nr.employed' to '#.employees'

colnames(bank)[colnames(bank) == "nr.employed"] <- "no.employees"

names(bank)
```

```{r View Atrribute Types}
sapply(bank,class)

#Convert appropriate attributes to "factor" type
bank <- as.data.frame(unclass(bank), stringsAsFactors = TRUE)

sapply(bank, class)
```
```{r Looking @ first six rows}
head(bank)
View(bank)
```
```{r Data Summary}
summary(bank)
```

```{r Null Value Check}
is.null(bank)
```



```{r Correlation Matrix}

numeric_data <- bank[, c("age", "cons.price.idx", "cons.conf.idx", "euribor3m", "no.employees" )]
correlation <- cor(numeric_data)

round(correlation,2)

##From the correlation matrix, we see that "emp.var.rate" and "euribor3m" have a strong correlation. We can go ahead and remove one of the columns.

bank <- subset(bank, select = -emp.var.rate)

names(bank)
sapply(bank, class)

##Checking Correlation Matrix

round(correlation,2)

```

---

The attribute “y” is binary and is our class attribute. The “yes” means the customer opened a term deposit and “no” means they did not. We will use this attribute to create our “Conversion Rate” KPI, by encoding 1 for “yes” and 0 for “no”.

```{r Encoding Atrribute}
bank <- bank %>%mutate(y=ifelse(y=="no", 0, 1))
bank$y <- as.integer(bank$y)
```


```{r Total conversions}
#Total Conversions
sum(bank$y)
```

```{r Total clients in the data}
#Total clients in the data
nrow(bank)
```

```{r Conversion rate}
#Conversion Rate
round(sum(bank$y)/nrow(bank)*100.0,2)
```
---
Our conversion rate came to 11.27%. Knowing this, we would like to assess the conversion rate against various attributes in this dataset.

### Age

```{r Conversion rate by Age}

#First, we create bins for the Age attribute (18-30, 30-40, 40-50, 50-60, 60-70, >70)
AgeConversions <- bank %>%
  group_by(AgeGroup=cut(age, breaks=seq(20, 70, by=10))) %>%
  summarize(TotalCount=n(), NumberConversions=sum(y)) %>%
  mutate(ConversionRate=NumberConversions/TotalCount*100)

#Rename the last bin
AgeConversions$AgeGroup <- as.character(AgeConversions$AgeGroup)
AgeConversions$AgeGroup[6] <- "70+"

#Create Histogram for Age Conversion Rate
ggplot(data=AgeConversions, aes(x=AgeGroup, y=ConversionRate)) +
  geom_bar(width=0.5, stat="identity", fill="darkblue") + 
  labs(title="Conversion Rates by Age")

```

What our histogram tells us is that the bank’s older customers are the ones who are responding well to the campaign.

### Age and Marital Status

```{r}

# Group data
AgeMaritalConversions <- bank %>%
  group_by(AgeGroup=cut(age, breaks=seq(20,70, by=10)),
           Marital=marital) %>%
  summarize(Count=n(), NumConversions=sum(y)) %>%
  mutate(TotalCount=sum(Count)) %>%
  mutate(ConversionRate=NumConversions/TotalCount*100)

#Group Name Change
AgeMaritalConversions$AgeGroup <- as.character(AgeMaritalConversions$AgeGroup)
AgeMaritalConversions$AgeGroup[is.na(AgeMaritalConversions$AgeGroup)] <- "70+"

#Pair Age and Marital Status and gauge conversion rate
ggplot(AgeMaritalConversions, aes(x=AgeGroup, y=ConversionRate, fill=Marital)) +
  geom_bar(width=0.5, stat = "identity") +
  labs(title="Conversion Rates by Age Group and Marital Status")

```
Married people are our highest converters. Considering the age groups where these people can be found, it makes sense. The nature of a term deposit product would best resonate with people in those demographics because it’s a more of a medium to long-term investment product.

### Job/Occupation

```{r}

 #Group Data

JobConversions <- bank %>%
  group_by(Job=job) %>%
  summarize(TotalCount=n(), NumberConversions=sum(y)) %>%
  mutate(ConversionRate=NumberConversions/TotalCount*100) %>%
  arrange(desc(ConversionRate))

#Order By Job (DESC)

JobConversions$Job <- factor(JobConversions$Job, 
                                   levels = JobConversions$Job[order(-JobConversions$ConversionRate)])

# Plotting conversions by Job

ggplot(JobConversions, aes(x=Job, y=ConversionRate)) +
  geom_bar(width=0.5, stat = "identity", fill="darkred") +
  labs(title="Conversion Rates by Job") +
  theme(axis.text.x = element_text(angle = 90))

```
The plot suggest students and retired persons are the highest converters. The high conversion rate makes sense, as it matches with the Age plot we saw previously. More investigation may be necessary as to why the conversion rate for students is so high, considering the conversion rates in other attributes/demographics.

### Account in Default

```{r}
#Create Object

DefaultCredit <- bank %>%
  group_by(DefaultState=default) %>%
  summarize(TotalCount=n(), NumberConversions=sum(y)) %>%
  mutate(ConversionRate=NumberConversions/TotalCount*100) %>%
  arrange(desc(ConversionRate))

#Visualizing conversion rate

ggplot(DefaultCredit, aes(x=DefaultState, y=ConversionRate, fill=DefaultState)) +
  geom_bar(width=0.5, stat = "identity") +
  labs(title="Conversion Rates by Default Credit")

```
This outcome was expected. Those customers who do not have a credit account in default are high converters. Based on our findings, there would be no need to focus marketing efforts on those that do.

### Call Duration

```{r}
#Create Object

call_duration <- bank %>%
  group_by(Subscribed=y) %>%
  summarise(Average_Duration=mean(duration))

head(call_duration)

```

This tells us that it takes longer calls, on average, to get people to convert. Could there be a way to shorten call time and still convert? Further research would need to be done. Could there be certain keywords or phrases that occur during conversion calls? Could a “conversion script” be created for agents that increases conversion rate while reducing call time?

### Contact Type

```{r}
#Create Object

ContactType <- bank %>% group_by(Contact=contact) %>% summarize(TotalCount=n(), NumberConversions=sum(y)) %>% mutate(ConversionRate=NumberConversions/TotalCount*100) %>% arrange(desc(ConversionRate))

head(ContactType)

```
#### Conversion Month

```{r}

# Create Object

MonthConvert <- bank %>% group_by(Month= month) %>% summarize(TotalCount=n(), NumberConversions=sum(y)) %>% mutate(ConversionRate=NumberConversions/TotalCount*100) %>% arrange(desc(ConversionRate))

#Sort in DESC
MonthConvert$Month <- factor(MonthConvert$Month, 
                                   levels = MonthConvert$Month[order(-MonthConvert$ConversionRate)])
#Create Viz
ggplot(MonthConvert, aes(x=Month, y=ConversionRate)) +
  geom_bar(width=0.5, stat = "identity", fill="darkorange") +
  labs(title="Conversion Rates by Month Contacted") +
  theme(axis.text.x = element_text(angle = 90))
```

##Conversion by Day of Week
```{r}

#Create Object
DayOfWeek <- bank %>%
  group_by(Day_Of_Week=day_of_week) %>%
  summarize(TotalCount=n(), NumberConversions=sum(y)) %>%
  mutate(ConversionRate=NumberConversions/TotalCount*100) %>%
  arrange(desc(ConversionRate))

#Sort DESC

DayOfWeek$Day_Of_Week <- factor(DayOfWeek$Day_Of_Week, levels = c("mon", "tue", "wed", "thu", "fri"))
#Create Viz
ggplot(DayOfWeek, aes(x=Day_Of_Week, y=ConversionRate)) +
  geom_bar(width=0.5, stat = "identity", fill="darkblue") +
  labs(title="Conversion Rates by Day of Week Contacted") +
  theme(axis.text.x = element_text(angle = 90))
```

We can see that people are more likely to convert on Thursdays, Tuesdays and Wednesdays in March, December, September and October.
```{r}
str(bank)
```
```{r}
bank['month_num'] = as.numeric(bank$month)
bank['job_num'] = as.numeric(bank$job)
bank['marital_num'] = as.numeric(bank$marital)
bank['education_num'] = as.numeric(bank$education)
bank['default_num'] = as.numeric(bank$default)
bank['housing_num'] = as.numeric(bank$housing)
bank['loan_num'] = as.numeric(bank$loan)
bank['contact_num'] = as.numeric(bank$contact)
bank['day_of_week_num'] = as.numeric(bank$day_of_week)
bank['poutcome_num'] = as.numeric(bank$poutcome)


```
```{r}
str(bank)
```
```{r}
bank_sub <- subset(bank, select = -c(2:10,15))
bank_sub
str(bank_sub)
```
```{r}
round(cor(bank_sub),2)
```
```{r}
pc_bank_sub <- princomp(bank_sub,cor = TRUE , scores = TRUE)
pc_bank_sub

summary(pc_bank_sub)
```
```{r}
plot(pc_bank_sub, col="Blue")

```

```{r}
pc_bank_sub_scores <- pc_bank_sub$scores
head(pc_bank_sub_scores[,c(1:13)])
```
```{r}
nbank<- pc_bank_sub_scores
colnames(pc_bank_sub_scores)
```




```{r}
nbank_ctree<- ctree(y ~., data = nbank)


plot(nbank_ctree, type = "simple" )


```
```{r}
train_index<- sample (1:nrow(bank),0.7* nrow(bank))
train.set<- bank[train_index,]
```


```{r}
test.set <- bank[-train_index,]
```


```{r}
bank_ctree_model <- ctree( y~.,data= train.set)
bank_ctree_model
```
```{r}
bank_ctree_prediction <- predict(bank_ctree_model, test.set)
head(bank_ctree_prediction)
```
```{r}
table(bank_ctree_prediction, test.set$y)
```
```{r}
train_set <- bank[train_index,]
head(train_set)
```
```{r}
test_set = bank[-train_index,]
```

```{r}
indep_train_set <- train_set[-20]
indep_train_set
```
```{r}
indep_test_set<- train_set[-20]
indep_test_set
```
```{r}
target_train_set <- train_set$y
target_train_set
```
```{r}
target_test_set <- test_set$y
target_test_set
```
```{r}
pred_target_test_set <- knn(indep_train_set, indep_test_set ,target_train_set,k=3)
pred_target_test_set
```
```{r}
table(pred_target_test_set, target_test_set)
```





### Initial Findings

From this initial analysis, we've been able to glean some useful insights into the bank's customers and would be able to create a working profile of "ideal" customers. As the project evolves and more analysis is conducted, we should be able to derive more insights and possibly build useful models.